---
title: "R Notebook-morphoMS"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
  word_document: default
---


try this  things again
```{r load problematic packages}
install.packages("devtools")
install.packages("ggplot2")
install.packages("ggpubr")
install.packages("factoextra")
install.packages("FactoMineR")
install.packages("ade4")
install.packages("vegan")
install.packages("usethis")
install.packages("ggsignif")
install.packages("RColorBrewer")
library(devtools)
library(ggplot2)
library(ggpubr)
library(factoextra)
library(FactoMineR)
library(RColorBrewer)
```


```{r load dataset}
allrhino58traits = read.csv(file = "rhinoall58traits.csv", header = T, sep = ",")
allrhino58traits
str(allrhino58traits)
allrhino58traits$Id
```

subset

```{r SUBSET DATA}
#subset and do summary per species
#affinis
aff = subset(allrhino58traits, Species=="Rhinolophus_affinis")
aff
str(aff)
sum.aff = summary(aff)
sum.aff

#luctus
luc = subset(allrhino58traits, Species == "Rhinolophus_luctus")
luc
str(luc)
sum.luc = summary(luc)
sum.luc

#malayanus
mal = subset(allrhino58traits, Species == "Rhinolophus_malayanus")
mal
str(mal)
sum.mal = summary(mal)
sum.mal

#marshalli
mar = subset(allrhino58traits, Species == "Rhinolophus_marshalli")
mar
str(mar)
sum.mar = summary(mar)
sum.mar

#paradoxolophus
par = subset(allrhino58traits, Species == "Rhinolophus_paradoxolophus")
par
str(par)
sum.par = summary(par)
sum.par

#pearsonii
pea = subset(allrhino58traits, Species == "Rhinolophus_pearsonii")
pea
str(pea)
sum.pea = summary(pea)
sum.pea

#pusillus
pus = subset(allrhino58traits, Species == "Rhinolophus_pusillus")
pus
str(pus)
sum.pus = summary(pus)
sum.pus

#macrotis
mac = subset(allrhino58traits, Species == "Rhinolophus_siamensis")
str(mac)
sum.mac = summary(mac)
sum.mac

#sinicus
sin = subset(allrhino58traits, Species == "Rhinolophus_sinicus")
sin
str(sin)
sum.sin = summary(sin)
sum.sin

#stheno
sth = subset(allrhino58traits, Species =="Rhinolophus_stheno")
str(sth)
sum.sth = summary(sth)
sum.sth
```

look if FA correlate with fmax
```{r aimple linear regression}
library(car)

#homogeneity
leveneTest(FA ~ sp_id, data = allrhino58traits)
# Compute the analysis of variance
FA.aov <- aov(FA  ~ sp_id, data = allrhino58traits)
# 2. Normality
plot(FA.aov, 2) #qqplot
plot(FA.aov, 1) #residual vs fitted plot
# Summary of the analysis
summary(FA.aov) #result are significant
#do tukey (posthoc)
TukeyHSD(FA.aov)
#Extract the residuals
FA.aov_residuals <- residuals(object = FA.aov )
# Run Shapiro-Wilk test
shapiro.test(x = FA.aov_residuals)
#do alternative non parametric kruskal wallis
kruskal.test(FA~ sp_id, data = allrhino58traits)

```

do linear regression

1. y is fmaxe (if calls are correlated with body size parameters)

linear regression assumption
Linear regression makes several assumptions about the data, such as :

Linearity of the data. The relationship between the predictor (x) and the outcome (y) is assumed to be linear.
Normality of residuals. The residual errors are assumed to be normally distributed.
Homogeneity of residuals variance. The residuals are assumed to have a constant variance (homoscedasticity)
Independence of residuals error terms.
You should check whether or not these assumptions hold true. Potential problems include:

Non-linearity of the outcome - predictor relationships
Heteroscedasticity: Non-constant variance of error terms.
Presence of influential values in the data that can be:
Outliers: extreme values in the outcome (y) variable
High-leverage points: extreme values in the predictors (x) variable
All these assumptions and potential problems can be checked by producing some diagnostic plots visualizing the residual errors.

```{r regression AND LINEAR MODEL}

#look at the relationship between fa and fmaxe
library(tidyverse)
library(ggpubr)
#first, plot data against predictor (FmaxE), we assume body size related to fmaxe
ggplot(allrhino58traits, aes(x = NL.L , y = FmaxE)) + geom_point() +stat_smooth()
#correlation coefficien
cor(allrhino58traits$FmaxE, allrhino58traits$NL.L)
#do linear model to determine a and b (linear regression, y = a+bx)
NL.Lmodel = lm(FA ~ B, data = allrhino58traits)
NL.Lmodel

#add regression line in the scatter plot
ggplot(allrhino58traits, aes(NL.L, FmaxE)) +
  geom_point() +
  stat_smooth(method = lm)
ggsave("NL.LvsFmax.tiff", units="in", width=6, height=6, dpi=300, compression = 'lzw')
#do model summary
summary(NL.Lmodel)

#95% confidence interval
confint(NL.Wmodel)

#model accuracy
sigma(NL.Wmodel)*100/mean(allrhino58traits$NL.W)


#add to the formula and R to graph
eq <- function(x,y) {
  
  model <- lm(FmaxE ~ INCW, data = allrhino58traits)
  as.character(
    as.expression(
      substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                list(a = format(coef(model)[1], digits = 4),
                b = format(coef(model)[2], digits = 4),
                r2 = format(summary(model)$r.squared, digits = 3)))
    )
  )
}

ggplot(allrhino58traits,aes(x = INCW, y = FmaxE)) + 
  geom_point() + 
  geom_smooth(method = lm) +
  geom_text(x = 7, y = 100, label = eq(allrhino58traits$INCW, allrhino58traits$FmaxE), parse = TRUE) 

summary(model)
ggsave("FmaxEVSINCW.tiff", units="in", width=6, height=6, dpi=300, compression = 'lzw')

#check the regression
par(mfrow = c (2,2))
plot(famodel)
install.packages("ggfortify")
library(ggfortify)
autoplot(famodel)
```

#do multiple regression 

```{r multiple regression}
eq <- function(x,y) {
  
  multiplemodel <- lm(FmaxE ~ E*S.I, data = allrhino58traits)
  as.character(
    as.expression(
      substitute(italic(y) == a + b %.% italic(x)*","~~italic(r)^2~"="~r2,
                list(a = format(coef(model)[1], digits = 4),
                b = format(coef(model)[2], digits = 4),
                r2 = format(summary(model)$r.squared, digits = 3)))
    )
  )
}
summary(multiplemodel)
equation1=function(x){coef(multiplemodel)[2]*x+coef(multiplemodel)[1]}
equation2=function(x){coef(multiplemodel)[2]*x+coef(multiplemodel)[1]+coef(multiplemodel)[3]}

ggplot(allrhino58traits,aes(y=FmaxE,x=E,color=S.I))+geom_point()+
        stat_function(fun=equation1,geom="line",color=scales::hue_pal()(2)[1])+
        stat_function(fun=equation2,geom="line",color=scales::hue_pal()(2)[2])

summary(model)
ggsave("FmaxEVSINCW.tiff", units="in", width=6, height=6, dpi=300, compression = 'lzw')
```

#do boxplot

```{r tralala}

pdf("macrotis CP.S.Ang.pdf")
ggplot(mac, aes(x=sp_id, y=CP.S.Ang, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="blue", size=0.4, alpha=0.9) +
  theme_classic()+
  #theme_bw()+
  #scale_fill_manual(values = mycolors) +
  ggtitle("R. macrotis-CP.S.Ang") +
  xlab("Species") + ylab("CP.S.Ang")
dev.off()
ggsave("macrotis CP.S.Ang.tiff", units="in", width=8, height=6, dpi=300, compression = 'lzw')
```
#continue boxplot
```{r boxplot FA}
#do boxplot for FA all species
ggplot(allrhino58traits, aes(x=Species, y=FA, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("FA boxplot") +
  xlab("Species") + ylab("FA")
ggsave("FA.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


# Define the number of colors you want
nb.cols <- 18
mycolors <- colorRampPalette(brewer.pal(8, "Paired"))(nb.cols)
# Create a ggplot with 18 colors 
# Use scale_fill_manual
#ggplot(df) + 
  #geom_col(aes(name, Sepal.Length, fill = factor(Sepal.Length))) +
  #scale_fill_manual(values = mycolors) +
  #theme_minimal() +
  #theme(legend.position = "top") #use this as reference in the future

pdf("FAboxplot.pdf")
ggplot(allrhino58traits, aes(x=sp_id, y=FA, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("FA boxplot") +
  xlab("Species") + ylab("FA")
dev.off()
ggsave("FA.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```

#do boxplot for Ears
```{r boxplot E}
ggplot(allrhino58traits, aes(x=Species, y=E, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("E boxplot") +
  xlab("Species") + ylab("E")
ggsave("E.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')

pdf("EarsBoxplot.pdf")
ggplot(allrhino58traits, aes(x=sp_id, y=E, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("E boxplot") +
  xlab("Species") + ylab("E")
dev.off()
ggsave("E.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```

#DO BOXPLOT FOR TAIL---

```{r BOXPLOT FOR TAIL}
ggplot(allrhino58traits, aes(x=Species, y=Tail, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Tail boxplot") +
  xlab("Species") + ylab("Tail")
ggsave("Tail.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')

pdf("tailboxplot.pdf")
ggplot(allrhino58traits, aes(x=sp_id, y=Tail, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("Tail boxplot") +
  xlab("Species") + ylab("Tail")
dev.off()
ggsave("Tail.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#boxplot for FmaxE

```{r BOXPLOT FOR FmaxE}
#do boxplot for FmaxE all species
ggplot(allrhino58traits, aes(x=Species, y=FmaxE, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("FmaxE boxplot") +
  xlab("Species") + ylab("FmaxE")
ggsave("FmaxE.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')

pdf("FMAXEboxplot.pdf")
ggplot(allrhino58traits, aes(x=sp_id, y=FmaxE, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("FmaxE boxplot") +
  xlab("Species") + ylab("FmaxE")
dev.off()
ggsave("FmaxE.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```

#do boxplot for BW-BANDWIDTH
```{r boxplot BW}
ggplot(allrhino58traits, aes(x=Species, y=BW, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("BW boxplot") +
  xlab("Species") + ylab("BW")
ggsave("BW.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')

pdf("BWboxplot.pdf")
ggplot(allrhino58traits, aes(x=sp_id, y=BW, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("BW boxplot") +
  xlab("Species") + ylab("BW")
dev.off()
ggsave("BW.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#do boxplot for wing
# boxplot wor wing span (B)
```{r boxplot for wingspan b}
ggplot(allrhino58traits, aes(x=Species, y=B, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("B boxplot") +
  xlab("Species") + ylab("B")
ggsave("B.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')

pdf("wingspanB.pdf")
ggplot(allrhino58traits, aes(x=sp_id, y=B, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("B boxplot") +
  xlab("Species") + ylab("B")
dev.off()
ggsave("B.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```

```{r BOXPLOT FOR wing tip shape I}
ggplot(allrhino58traits, aes(x=Species, y=I, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("I boxplot") +
  xlab("Species") + ylab("I")
ggsave("I.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=I, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("I boxplot") +
  xlab("Species") + ylab("I")
ggsave("I.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#do boxplot for ASPECT RATIO

```{r boxplot for AR}

ggplot(allrhino58traits, aes(x=Species, y=AR, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("AR boxplot") +
  xlab("Species") + ylab("AR")
ggsave("AR.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=AR, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("AR boxplot") +
  xlab("Species") + ylab("AR")
ggsave("AR.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#DO BOXPLOT FOR WINGLOADING

```{r BOXPLOT WL}
ggplot(allrhino58traits, aes(x=Species, y=WL, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("WL boxplot") +
  xlab("Species") + ylab("WL")
ggsave("WL.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=WL, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("WL boxplot") +
  xlab("Species") + ylab("WL")
ggsave("WL.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#DO BOXPLOT FOR BODY MASS

```{r BOXPLOT FOR BODYMASS}
ggplot(allrhino58traits, aes(x=Species, y=BM, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("BM boxplot") +
  xlab("Species") + ylab("BM")
ggsave("BM.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=BM, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("BM boxplot") +
  xlab("Species") + ylab("BM")
ggsave("BM.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```

##DO BOXPLOT FOR NOSELEAF TRAITS

```{r DO BOXPLOT FOR NLL}
ggplot(allrhino58traits, aes(x=Species, y=NL.L, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("NLL boxplot") +
  xlab("Species") + ylab("NLL")
ggsave("NLL.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')

pdf("NLLboxplot.pdf")
ggplot(allrhino58traits, aes(x=sp_id, y=NL.L, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("NLL boxplot") +
  xlab("Species") + ylab("NLL")
dev.off()
ggsave("NLL.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#DO BOXPLOT FOR NLW

```{r BOXPLOT NLW}
ggplot(allrhino58traits, aes(x=Species, y=NL.W, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("NL.W boxplot") +
  xlab("Species") + ylab("NL.W")
ggsave("NL.W.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=NL.W, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("NL.W boxplot") +
  xlab("Species") + ylab("NL.W")
ggsave("NL.W.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#DO BOXPLOT FOR LH

```{r BOXPLOT FOR LH}
ggplot(allrhino58traits, aes(x=Species, y=LH, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("LH boxplot") +
  xlab("Species") + ylab("LH")
ggsave("LH.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=LH, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("LH boxplot") +
  xlab("Species") + ylab("LH")
ggsave("LH.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#DO BOXPLOT FOR LTH

```{r BOXPLOT FOR LTH}
ggplot(allrhino58traits, aes(x=Species, y=LTH, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("LTH boxplot") +
  xlab("Species") + ylab("LTH")
ggsave("LTH.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=LTH, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("LTH boxplot") +
  xlab("Species") + ylab("LTH")
ggsave("LTH.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```

#DO BOXPLOT FOR NOSELEAF AREA

```{r BOXPLOT FOR NL.A}
ggplot(allrhino58traits, aes(x=Species, y=NL.A, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("NL.A boxplot") +
  xlab("Species") + ylab("NL.A")
ggsave("NLA.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=NL.A, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("NL.A boxplot") +
  xlab("Species") + ylab("NL.A")
ggsave("NL.A.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#DO BOXPLOT FOR SELLA HEIGHT

```{r BOXPLOT OF SELLA HEIGHT}
ggplot(allrhino58traits, aes(x=Species, y=Sh, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Sh boxplot") +
  xlab("Species") + ylab("Sh")
ggsave("Sh.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=Sh, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("Sh boxplot") +
  xlab("Species") + ylab("Sh")
ggsave("Sh.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#do boxplot for sella area

```{r Box[;ot for sella area]}
ggplot(allrhino58traits, aes(x=Species, y=SA, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("SA boxplot") +
  xlab("Species") + ylab("SA")
ggsave("SA.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=SA, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("SA boxplot") +
  xlab("Species") + ylab("SA")
ggsave("SA.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#DO BOXPLOT FOR ANGLE

```{r BOXPLOT CPANG}
ggplot(allrhino58traits, aes(x=Species, y=CP.Ang, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("CP.Ang boxplot") +
  xlab("Species") + ylab("CP.Ang")
ggsave("CP.Ang.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')

pdf("CPAngleboxplot.pdf")
ggplot(allrhino58traits, aes(x=sp_id, y=CP.Ang, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("CP.Ang boxplot") +
  xlab("Species") + ylab("CP.Ang")
dev.off()
ggsave("CP.Ang.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```
#do boxplot for CP.S.ang

```{r boxplot for CP.S.Ang}
ggplot(allrhino58traits, aes(x=Species, y=CP.S.Ang, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("CP.S.Ang boxplot") +
  xlab("Species") + ylab("CP.S.Ang")
ggsave("CP.S.Ang.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=CP.S.Ang, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("CP.S.Ang boxplot") +
  xlab("Species") + ylab("CP.S.Ang")
ggsave("CP.S.Ang.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```

#DO BOXPLOT FOR S.ANG

```{r BOXPLOT FOR S.ANG}
ggplot(allrhino58traits, aes(x=Species, y=S.Ang, fill=Species)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_brewer(palette = "Paired") +
  ggtitle("CP.Ang boxplot") +
  xlab("Species") + ylab("S.Ang")
ggsave("S.Ang.boxplot.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')


ggplot(allrhino58traits, aes(x=sp_id, y=S.Ang, fill=sp_id)) +
  geom_boxplot(color="black", alpha=0.3) +
  geom_jitter(color="red", size=0.4, alpha=0.9) +
  theme_classic()+
  scale_fill_manual(values = mycolors) +
  ggtitle("S.Ang boxplot") +
  xlab("Species") + ylab("S.Ang")
ggsave("S.Ang.boxplot.species.tiff", units="in", width=10, height=6, dpi=300, compression = 'lzw')
```

















